On the properties of discs around accreting brown dwarfs

### 2010-05-24 BD_disc/paper.tex
Completely changed results section and conclusions-these sections need finessing
git-svn-id: https://repository.astro.ex.ac.uk/torus/trunk@2281 25464f21-6099-402c-b30a-640f1d40d38b

\begin{abstract}
To Do
\begin{enumerate}
\item{Read new papers and put in any relevant stuff.}
\item{Sort structure paragraph at end of intro.}
\item{Sort out headings of sections.}
\item{Sort clearpage problem, perhaps change to includegraphics.}
\item{Sort out the webpage stuff.}
\end{enumerate}
\end{abstract}

@@

In surveys, particularly those attempting to discern the PMS disc
fraction , this disentangling usually takes place using broad-band
photometric measures and cuts in colour-colour or colour-magnitude
space
\citep{luhman_2005,Luhman_2008,gutermuth_2008}.
The important quantity here is the separation in wavelength between
the emission peaks for the stellar and thermal disc components, as it
is these components which must be separated. For BDD systems it is
likely that difficulties detecting the disc will be exacerbated by the
lower temperatures of the photosphere and therefore the smaller
separation in wavelength from the thermal disc emission component,
compared to CTTS systems. Therefore, to derive disc fractions for BD
populations we require detailed comparison models with which to guide
disc candidate selection.

@@

Here we present a grid of models of BDD systems, including a
self-consistent treatment of the photospheric and accretion luminosity
sources and the interaction of that flux with the circumstellar disc.
We investigate the impact of the luminosity of the central source on
location and shape of the disc inner rim, as well as the large-scale
structure and flaring of the outer disc. We construct synthetic
colour-colour and colour-magnitude diagrams in order to examine the
efficacy of the photometric selections used to isolate brown dwarfs
and measure disc fractions
\citep{luhman_2005,Luhman_2008,gutermuth_2008}.

@@

Once the radiative transfer code was completed simulated SEDs were
generated. These SEDs can be generated for any distance and for any
system-observer inclination. For our models we have set the distance,
to 10\,pc to create absolute flux SEDs and selected ten inclinations
equi-spaced in $\cos i$ (0, 27, 39, 48, 56, 64, 71, 77, 84 and
90$^{\circ}$). A further useful feature of the {\sc torus} code is
that the emitted photon packets which make up the SED are tagged on their way
to the observer. These tags separate the packets into four
groups. Firstly, packets are separated by source into thermal (disc)
or stellar groups. These groups are then subdivided into those which
reach the observer either directly or after scattering (for example
see Figure \ref{disc_eg_173}). The resulting SEDs are discussed in
Section \ref{seds}.

@@

These specific filter sets have been chosen due to their ubiquitous
use and suitability for the derivation of the key stellar parameters
of age, mass and, for populations, disc fractions. Therefore, by
studying the changes of magnitudes (and colours) in these photometric
systems we can test the predicted effects on these derived parameters
caused by changes in the input parameters of our model grid. The
optical magnitudes \textit{VI} are used to explore age related effects
of varying the parameter space. Flux in the \textit{VI} bands is only
minimally affected by accretion flux \citep{gullbring_1998} and
disc thermal flux \citep{hartmann_1998} and, additionally, in a
\textit{V, V-I} CMD, the reddening vector lies parallel to the pre-MS,
minimising any age effect of extinction uncertainty, \citep[for a full
discussion see][]{mayne_2008,mayne_2007}. The
near-IR passbands of \textit{JHK} are most often used to derive
stellar masses as for pre-MS objects the reddening vector, in for
instance, a \textit{J, J-K} CMD, is almost perpendicular to the
isochrones, minimising the mass effect of extinction uncertainties
\citep[the \textit{CIT} systems was chosen to
match][]{chabrier_2000}. Finally, as is now well documented the
\textit{Spitzer} IRAC passbands provide the best data with which to
unambiguously separate naked and star-disc systems
\citep{luhman_2005}. In addition, at longer wavelengths, the
MIPS instrument provides sensitivity to disc systems at much larger
radii (or debris discs).

@@

In this Section we first discuss the physical structure, both density
and temperature, of the BDD disc systems (Section \ref{disc_struct})
across our parameter space. Then we discuss the resulting simulated
SEDs for these BDD systems (Section \ref{observables}). Finally, we
produce simulated photometric observations (Section \ref{photometry}).
These simulated observations allows us to show that recovery of the
input masses and ages for our accreting BDD systems is unreliable. We
show that despite an intrinsic relationship between rotation rate and
initial inner edge position we do not recover a rotation rate
correlation with IR excess. In addition, despite not intrinsically
including a dependence of accretion rate on stellar mass in our grid.
We show that current observational techniques and theoretical models
applied to the grid would result in a relationship of this type being
derived.

@@

The initial shape of the inner edge of the dust disc is a vertical
wall coincident with the co-rotation radius. In the previous section
we have shown that models with a negligible accretion rate do not
significantly sublimate the dust, and hence the edge remains
vertical. This inner edge is heated by direct radiation from the
protostar, and its scaleheight increases. Disc material behind the
inner rim is shielded from direct radiation and has a smaller
scaleheight, leading to the `puffed up' inner rim predicted by
\cite{dullemond_2001}. This effect is illustrated in
Figure~\ref{no_sub_183}, a model in which there is negligible dust
sublimation.

@@

@@

The left panels show accretion streams with an areal coverage of 10\%
and the left panels 1\%. As the areal coverage reduces the effective
temperature of the accretion hot spot increases, resulting in an
increase in accretion flux, and resulting shift to bluer wavelengths
of the peak flux. This can be seen clearly by comparing the left and
right panels of Figure \ref{acc_dom}. Perhaps the most important,
albeit qualitative, result shown is Figure \ref{acc_dom} is an insight
into the accretion rate at which the accretion blackbody flux
dominates over the photospheric flux. Figure \ref{acc_dom} shows that
as the accretion rate raises above \logmdot = $-9$ for systems with 1
or 10\% areal coverage, the accretion flux dominates the emergent SED
at both periods. Effectively, the spectroscopic features of the
photosphere are veiled by the additional continuum accretion
flux. Therefore for reasonable coverages (1--10\%) and rotational
periods (0.5--5 days) the photospheric flux is effectively veiled by
accretion flux for accretion rates \logmdot$>-9$. It is also clear
from Figure \ref{acc_dom} that the magnitude becomes brighter and the
colour bluer (in terms of optical photometry) with increasing
accretion rates as expected \citep{gullbring_1998}.

@@

As shown in Figures \ref{flare_-12_183} and \ref{flare_-7_193} and
discussed in Section \ref{disc_flaring} BDD systems under vertical
hydrostatic equilibrium have highly flared discs. As discussed in
\cite{walker_2004} this increased flaring (when compared to CTTS
stars) leads to occultation at the star at lower inclination angles
and, therefore, significant changes to the SED. Increases in the
inclination angle, for these systems, quickly lead a significant
proportion of the stellar flux being intercepted, and reprocessed, by
the highly flared disc. This reprocessing will lead to a change in the
flux levels at the shorter, bluer, wavelengths as more stellar flux is
intercepted by the disc. It will also lead to significant changes in
the flux reaching the observer from the inner and outer regions of the
disc as the system approaches edge on. Also, as discussed in
Section \ref{disc_struct} the addition of dust sublimation leads to a
change in the shape of the inner edge, where the temperature is
sufficient. Figures \ref{no_sub_183} to \ref{sub_179} \ref{sub_181}
show a range in inner edge shapes, from flat walls through concave to
convex curves, caused by the radial density profile of the disc and
the dependence of the sublimation temperature on density. This change
in shape, as noted for Herbig Ae stars by \cite{tannirkulam_2007}
XX CHECK FOR A CTTS REFERENCESXXX, will lead to changes in the
characteristics of the SED, or derived IR excess, with inclination
angle \citep{tannirkulam_2007}, for close to face on viewing
angles.

@@

HERER****
 write caption
GOT HERE*** TUNE WHAT NEED FROM THIS FIGURE***

SELECT THE CORRECT MODELS

POINTS TO MAKE, MJ (and MV) brighter for high accretion, tail off
starts earlier for high accretion

J-K delta larger over short inclinations than low accretors 0.05 0.1
factor two larger in magnitude!!! factor ten in flux?

The changes in flux with inclination will clearly lead to changes in
the observed magnitudes with inclination. Figure \ref{flare_inner_mag}
shows the $M_V$ and $M_J$ magnitude, in the \textit{top} and
\textit{bottom panels} respectively, as a function of inclination with
the individual points marked as crosses. The \textit{top} and
\textit{bottom panels} show accretion rates of -12, -9 and -7
log$(\frac{\dot{M}}{M_{\odot}}yr^{-1})$ as black, red and blue lines.
The solid lines are for objects with rotational periods of 0.5 days
and the dashed lines are for periods of 5 days. The \textit{inset
  panel} within the \textit{bottom panels} shows an enlargement of the
section of inclinations from 40 to 80$^{\circ}$. For \textit{all
  panels} the vertical dotted lines show the inclinations for which
significant drops in flux are observed (see Figure \ref{flare_inc})
for -12 and -7 log$(\frac{\dot{M}}{M_{\odot}}yr^{-1})$ in black and
blue respectively.

\begin{figure*}
  \centering
  \includegraphics[scale=0.6,angle=90]{./Fig/flare_inner_mag.ps}
  \caption{Figure showing both XXX.\label{flare_inner_mag}}
\end{figure*}

XXXSUMMARISE THISXXX

@@

XXXADD LINE ABOUT OBJECT

@@

@@

For the derivation of ages optical CMDs, in particular in \textit{V,
  V-I}, are most often used, and indeed most suitable. Whereas, IR
CMDs, such as a \textit{J, J-K} CMD, are most suitable for mass
derivation \cite[see references and discussions in][ and Section
\ref{derived}]{2007MNRAS.375.1220M,2008MNRAS.386..261M}.

As discussed in Section \ref{sed_analysis} the presence of a
circumstellar disc and accretion result in significant changes to the
simulated SEDs for our models. Firstly, the photophere of the central
star is dominated by the accretion luminosities for accretion rates of
$\dot{M}>$10$^{-9}M_{\odot}yr^{-1}$ as shown in Section
\ref{acc_dominance}. Secondly, for increasing accretion rates the
outer disc flaring increases, resulting in occultation of the star at
smaller inclination angles, as discussed in Section
\ref{flare_obscure}. Finally, the location of the inner edge and flux
intercepted by this edge lead to changes in its shape and subsequent
dependence of the observed SED with inclination, as discussed in
Section \ref{inner_shape}. In this Section we explore the effects of
these changes in SED shape on the derived photometry in CMDs used for
age and mass derivation for stars. The use of isochrones for the
derivation of single star parameters is completely unreliable
\citep[see discussion in][]{2008MNRAS.386..261M}. Practically,
therefore, median ages are derived from populations. Subsequently,
derived masses are still unreliable but at least based on a consistent
age. This problem is being addressed by Bell et al (in prep), where
$K$ band photometry and known eclipsing binaries are being used to
refine pre-MS isochrones. In this section we plot the data for our 1
Myr systems only and explore the resulting scatters caused by the disc
presence and accretion luminosity.

As can be seen in Figures \ref{age_mass_VVI} and \ref{age_mass_JHK}
current accretion in a star and disc systems, for rates typical in the
BD mass regime, creates a scatter in our simulated photometry
indicative of a much larger isochronal age spread than 10 Myr. Indeed,
for many BDD systems even at nominal accretion rates of -11 or -12
\logmdot, for our simulations, the colours of these stars move
significantly blueward of the expected BD locus in a $M_V$, $(V-I)_0$
CMD. As the dots in these Figures are the simulated photometry of BDD
systems over a range of typical input parameters (see Section
\ref{par_space}), an observed coeval 1 Myr population could reasonably
be expected to show a similar scatter. The naked isochrones show that
this spread in simulated photometry, would lead to a spread in
isochronal age greater than the $\approx$10 Myr spread claimed for
higher mass stars in some star forming regions (for instance the
ONC). Furthermore, for the higher accretion rates of -9 or -10
\logmdot the movement of the star within the CMD will effectively move
the star into the contamination region expected for background CTTS or
MS stars at a $(V-I)_0$ of $\leq$1.5 and $M_V$ 12-10, and as such the
star would not be included in a photometrically selected BD
sample. The solid green line, in the \textit{top left panel}, showing
the 1 Myr isochrone of \cite{2000A&A...358..593S} at a distance of 250
pc and extinction of $A_V=2$ mags shows that the BDD systems with
higher accretion rates could easily be confused for a background CTTS
or MS population (as MS stars are simply less luminous at a roughly
constant \textit{V-I} than the CTTS counterparts). Indeed miss
classification of a BDD system as a CTTS system has already been
revealed in \cite{2003ApJ...582.1109W}. Furthermore, scatter, although
somewhat reduced, can be observed in the simulated photometry of the
negligibly accreting systems, for example in the \textit{top left
  panel} some red dots, accreting at -12 or -11 \logmdot (which will
be negligible compared to the photospheric flux, see Figure
\ref{acc_dom}), still scatter significantly. These objects may still
be included in a `wide' photometric selection. This scatter is a
strong function of inclination, as shown in the \textit{bottom right
  panel}, where, as the inclination is increased the objects are
pushed lower in the CMD. Indeed, for the edge on cases some objects
have magnitudes fainter than show in Figure \ref{spread_opt_VVI_typ}
($M_V\approx 20$). This is expected as the star becomes obscured by
the flared disc, interestingly for these systems the \textit{bottom
  right panel} shows that this occurs for inclinations above around
71$^{\circ}$ (as found in Figure \ref{flare_inc}) in most cases.
However, even for the lower inclination angles some objects have very
faint magnitudes, this is due to the disc flaring leading to a smaller
opening angle and is discussed in Section \ref{disc_struct}. The
\textit{bottom left panel} shows that for naked systems accretion
rates of -9 \logmdot lead to the stars being scattered into the region
occupied by background CTTS and contamination (the lower accretion
rates are all coincident with the -12 \logmdot for 1 Myr). Crucially,
the \textit{top left panel} shows that the scatter from the isochrone
is, generally, correlated with accretion rate. Effectively, as the
accretion rate increases the BDD system moves farther away from the
isochrone and is therefore less likely to be classified as a BDD
system and included in any target samples of such objects. The
\textit{top right panel} shows that the scatter, in $M_V$, in our
simulated photometry from the isochrones is significant for both areal
coverages. The \textit{bottom left panel}, shows significant scatter
in the simulated photometry for both adopted rotation periods. This
panel also shows that scatter in this diagram is not obviously
correlated with the rotation period and therefore inner edge position.
Overall, as found in \cite{2004MNRAS.351..607W}, the dominant
scattering effect for BD disc systems in an optical CMD appears to be
caused by accretion rate and inclination, and therefore obscuration
effects of the disc on the star. This suggests that for a given
photometric survey of BDD systems accreting at typical accretion rates
and with an expected range of inclinations (centred on around
60$^{\circ}$) one would expect to exclude a significant fraction of
these objects from any isochrone based selection.  XXX EXPLAIN
FIGURESXXX

Figure \ref{spread_opt_VVI_ext} shows that for extreme accretion rates
the BDD objects move significantly blueward and brighter than the
naked isochrone (the naked isochrone at -9
\logmdot is shown for
comparison). Comparison of the \textit{four panels} again show that,
in general, the scatter is chiefly correlated with accretion rate and
inclination. Essentially, Figure \ref{spread_opt_VVI_ext} shows that
for, admittedly extreme, accretion rates BDD systems will not be
included in a photometrically (or indeed spectroscopically, from
Section \ref{sed_analysis}) selected sample.

>>>>

\begin{abstract}
To Do
\begin{enumerate}
\item{Read new papers and put in any relevant stuff.}
\item{Sort structure paragraph at end of intro.}
\end{enumerate}
\end{abstract}

@@

In surveys, particularly those attempting to discern the PMS disc
fraction , this disentangling usually takes place using broad-band
photometric measures and cuts in colour-colour or colour-magnitude
space
\citep{luhman_2005b,luhman_2008,gutermuth_2008}.
The important quantity here is the separation in wavelength between
the emission peaks for the stellar and thermal disc components, as it
is these components which must be separated. For BDD systems it is
likely that difficulties detecting the disc will be exacerbated by the
lower temperatures of the photosphere and therefore the smaller
separation in wavelength from the thermal disc emission component,
compared to CTTS systems. Therefore, to derive disc fractions for BD
populations we require detailed comparison models with which to guide
disc candidate selection.

@@

Here we present a grid of models of BDD systems, including a
self-consistent treatment of the photospheric and accretion luminosity
sources and the interaction of that flux with the circumstellar disc.
We investigate the impact of the luminosity of the central source on
location and shape of the disc inner rim, as well as the large-scale
structure and flaring of the outer disc. We construct synthetic
colour-colour and colour-magnitude diagrams in order to examine the
efficacy of the photometric selections used to isolate brown dwarfs
and measure disc fractions
\citep{luhman_2005b,luhman_2008,gutermuth_2008}.

@@

Once the radiative transfer code was completed simulated SEDs were
generated. These SEDs can be generated for any distance and for any
system-observer inclination. For our models we have set the distance,
to 10\,pc to create absolute flux SEDs and selected ten inclinations
equi-spaced in $\cos i$ (0, 27, 39, 48, 56, 64, 71, 77, 84 and
90$^{\circ}$). A further useful feature of the {\sc torus} code is
that the emitted photon packets which make up the SED are tagged on
their way to the observer. These tags separate the packets into four
groups. Firstly, packets are separated by source into thermal (disc)
or stellar groups. These groups are then subdivided into those which
reach the observer either directly or after scattering. The resulting
SEDs are discussed in Section \ref{seds}.

@@

These specific filter sets have been chosen due to their ubiquitous
use and suitability for the derivation of the key stellar parameters
of age, mass and, for populations, disc fractions. Therefore, by
studying the changes of magnitudes (and colours) in these photometric
systems we can test the predicted effects on these derived parameters
caused by changes in the input parameters of our model grid. The
optical magnitudes \textit{VI} are used to explore age related effects
of varying the parameter space. Flux in the \textit{VI} bands is only
minimally affected by accretion flux \citep{gullbring_1998} and
disc thermal flux \citep{hartmann_1998} and, additionally, in a
\textit{V, V-I} CMD, the reddening vector lies parallel to the pre-MS,
minimising any age effect of extinction uncertainty, \citep[for a full
discussion see][]{mayne_2008,mayne_2007}. The
near-IR passbands of \textit{JHK} are most often used to derive
stellar masses as for pre-MS objects the reddening vector, in for
instance, a \textit{J, J-K} CMD, is almost perpendicular to the
isochrones, minimising the mass effect of extinction uncertainties
\citep[the \textit{CIT} systems was chosen to
match][]{chabrier_2000}. Finally, as is now well documented the
\textit{Spitzer} IRAC passbands provide the best data with which to
unambiguously separate naked and star-disc systems
\citep{luhman_2005a}. In addition, at longer wavelengths, the
MIPS instrument provides sensitivity to disc systems at much larger
radii (or debris discs).

@@

In this Section we first discuss the physical structure, both density
and temperature, of the BDD disc systems (Section \ref{disc_struct})
across our parameter space. Then we discuss the resulting simulated
observations in Section\ref{observables}. We present analysis in terms
of the impacts of the disc structure on the SEDs and colours and
magnitudes. Then in Section \ref{isochrones} we examine the
reliability of age, mass and disc fraction derivation when applied to
our model grid. In particular we discuss selection effects causing
higher accreting systems to be unlikely to be classified as BDD
systems. Essentially, despite not intrinsically including a dependence
of accretion rate on stellar mass in our grid.  We show that current
observational techniques and theoretical models applied to the grid
would result in a relationship of this type being derived.

@@

The initial shape of the inner edge of the dust disc is a vertical
wall coincident with the co-rotation radius. In the previous section
we have shown that models with a negligible accretion rate do not
significantly sublimate the dust, and hence the edge remains
vertical. This inner edge is heated by direct radiation from the
protostar, and its scaleheight increases. Disc material behind the
inner rim is shielded from direct radiation and has a smaller
scaleheight, leading to the `puffed up' inner rim predicted by
\cite{dullemond_2001}. This effect is illustrated in
Figure~\ref{no_sub_183_a}, a model in which there is negligible dust
sublimation.

@@

XXXNJM HAS RE-WRITTEN FROM HEREXXX

@@

The left panels show accretion streams with an areal coverage of 10\%
and the left panels 1\%. As the areal coverage reduces the effective
temperature of the accretion hot spot increases, resulting in an
increase in accretion flux, and resulting shift to bluer wavelengths
of the peak flux. This can be seen clearly by comparing the left and
right panels of Figure \ref{acc_dom} As the rotation rate increases
the co-rotation radius also increases, as shown in Equation
\ref{inner_eq}, $R_{\rm inner}\propto \tau^{\frac{2}{3}}$. Therefore,
as Equations \ref{Lacc} and \ref{Tacc}, show, $L_{\rm acc}\propto 1-
\frac{R_*}{R_{\rm inner}}$ and $T_{\rm acc}\propto (\frac{L_{\rm
    acc}}{A})^{\frac{1}{4}}$. Therefore, the temperature of the
accretion hot spot increases as the rotational period increases, due
to the increase in potential energy lost by the mass accreted. This
can be seen in Figure by comparing the left and right panels of Figure
\ref{acc_dom}. Perhaps the most important, albeit qualitative, result
shown is Figure \ref{acc_dom} is an insight into the accretion rate at
which the accretion blackbody flux dominates over the photospheric
flux. Figure \ref{acc_dom} shows that as the accretion rate raises
above \logmdot = $-9$ for systems with 1 or 10\% areal coverage, the
accretion flux dominates the emergent SED at both
periods. Effectively, the spectroscopic features of the photosphere
are veiled by the additional continuum accretion flux. Therefore for
reasonable coverages (1--10\%) and rotational periods (0.5--5 days)
the photospheric flux is effectively veiled by accretion flux for
accretion rates \logmdot$>-9$. It is also clear from Figure
\ref{acc_dom} that the magnitude becomes brighter and the colour bluer
(in terms of optical photometry) with increasing accretion rates as
expected \citep{gullbring_1998}.

@@

As shown in Figures \ref{flare_-12_183} and \ref{flare_-7_193} and
discussed in Section \ref{disc_flaring} BDD systems under vertical
hydrostatic equilibrium have highly flared discs. As discussed in
\cite{walker_2004} this increased flaring (when compared to CTTS
stars) leads to occultation at the star at lower inclination angles
and, therefore, significant changes to the SED. Increases in the
inclination angle, for these systems, quickly lead a significant
proportion of the stellar flux being intercepted, and reprocessed, by
the highly flared disc. This reprocessing will lead to a change in the
flux levels at the shorter, bluer, wavelengths as more stellar flux is
intercepted by the disc. It will also lead to significant changes in
the flux reaching the observer from the inner and outer regions of the
disc as the system approaches edge on. Also, as discussed in Section
\ref{disc_struct} the addition of dust sublimation leads to a change
in the shape of the inner edge, where the temperature is
sufficient. Figures \ref{no_sub_183_a} to \ref{sub_179_a}
\ref{sub_181_a} show a range in inner edge shapes, from flat walls
through concave to convex curves, caused by the radial density profile
of the disc and the dependence of the sublimation temperature on
density. This change in shape, as noted for Herbig Ae stars by
\cite{tannirkulam_2007} XX CHECK FOR A CTTS REFERENCESXXX, will lead
to changes in the characteristics of the SED, or derived IR excess,
with inclination angle \citep{tannirkulam_2007}, for close to face on
viewing angles.

@@

The changes in flux with inclination will clearly lead to changes in
the observed magnitudes and colours with inclination. Figure
\ref{flare_inner_mag} shows the $M_V$ and $M_J$ magnitudes and the
$(J-K)_0$ colour index, in the top, middle and bottom panels
respectively. The colours and magnitudes are then marked as crosses as
a function of inclination.  The systems shown in all panels have an
age of 1 Myr, mass of $0.04M_{\odot}$, rotation rates of 0.5 and an
areal coverage of 10\%. The systems with accretion rates of \logmdot
-12, -9 and -7 are shown in black, red and blue, respectively, with
the lowest panel omitting \logmdot =-9, and restricting the
inclination to viewing angles close to face-on.  The top and middle
panels also show the system with the highest accretion rate but with a
rotation rate of 5 days as a dashed line. The vertical dotted lines in
the lowest panel simply illustrate the total change in magnitude, over
the shown inclination range.

\begin{figure*}
  \centering
  \includegraphics[scale=0.6,angle=90]{./Fig/flare_inner_mag.ps}
  \caption{Figure showing $M_V$, $M_J$ magnitudes and $(J-K)_0$ colour
    index as a function of inclination, shown as top, middle and lower
    panels respectively. The magnitude and colours of systems with
    $M_*=0.04M_{\odot}$, an age of 1 Myr, areal coverage of 10\% and
    rotation rate of 0.5 days are shown as crosses. The systems with
    accretion rates of \logmdot = -12, -9 and -7, are shown in black,
    red and blue respectively. The lowest panel omits the \logmdot -9
    systems. The top and middle panels also the systems with the
    highest accretion rate but for a rotation rate of 5 days. The
    lowest panel also shows, as horizontal dotted lines, the total
    change of colour with the illustrated
    inclinations. \label{flare_inner_mag}}
\end{figure*}

Accretion flux was shown to dominate the underlying, intrinsic,
photospheric SED for accretion rates of \logmdot=-9 (see Figure
\ref{acc_dom}). As one would expect this leads to significant changes
in the derived photometric magnitudes for bands blueward of a few
microns, where the accretion and photospheric flux dominate. This is
shown in the top panel of Figure \ref{flare_inner_mag}. As the
accretion rate increases the system moves to brighter magnitudes in
$M_V$. The move to brighter magnitudes becomes significant once the
accretion rate exceeds \logmdot=-9. Again moving from the fast to slow
rotation period increases the potential energy released and in this
case results in a brighter $M_V$ magnitude. As one moves to longer
wavelength photometric bands, as shown in the middle panel. The
increase in accretion rate affects these magnitudes, i.e. $M_J$ are
les significantly affected. This is due to the significance of the
inner edge at these longer wavelengths. The lowest panel shows that as
the accretion rate is increase the IR colours generally get
redder. Again, this is caused by interaction with the disc, as more
flux is reprocessed by the inner edge.

The top and middle panels of Figure \ref{flare_inner_mag} shows the
change in magnitude caused by obscuration from the flared outer
disc. The occultation starts earlier for the higher accretion rates,
with the magnitudes in $M_V$ and $M_J$ moving fainter at earlier
inclinations for the higher accretion rates.

XXXNOT SURE ABOUT THIS NEXT BIT, PERHAPS A LITTLE WEAKXXX

The lowest panel of Figure \ref{flare_inner_mag} shows the difference
in the inclination dependence of the IR colour for large curved
(\logmdot=-7) and smaller vertical (\logmdot=-12) inner walls. The
$(J-K_0$ colour of the higher accretion rate system changes more as a
function of inclination when compared to the lower accretion rate
system. A total change in $(J-K)_0$ of $\approx$ 0.05 and 0.1 are
found for \logmdot=-12 and -7 respectively. This doubling in the
change in IR colour as a function of inclination for viewing angles
close to face-on, is caused by the larger (vertically) and convex
inner wall in the extreme accreting case. For the convex inner wall
the effective emitting area one observes changes as a function of
inclination, leading to this greater dependence of IR colour on
inclination.

Therefore, for individual systems changes in the accretion rate and
disc structure lead to significant changes in magnitudes and
colours. The data presented in this section are for an isolated mass
and age. However, the trends presented are present throughout our
model grid for any given subset, and as such are representative. The
scatter or changes in magnitude in the individual systems will change
as a function of the remaining variables, but the dominant input
variables affecting simulated photometry are accretion rate and
inclination. For populations of stars these changes in magnitude and
colour act to scatter or spread a pre-MS BD locus in photometric
space.

@@

@@

To delineate the effects of the accretion rate and circumstellar discs
we have subdivided the grid into two groups (as discussed in Sections
\ref{disc_struct}), those with accretion rates typical for higher mass
CTTS objects, defined as \logmdot=-12 (negligible) to \logmdot=-9 and
those with elevated accretion rates, where \logmdot>-9. For several of
the plots in this section the magnitude and colours for the
$M_*=0.01M_{\odot}$ systems at high inclinations become extremely
faint and red. In some cases these objects appear slightly without the
scale of the diagram the axes were limited in this way to be able to
show the changes in colour and magnitude for the majority of stars
better. The colours and magnitudes for these lowest mass stars are not
necessarily unreliable but simply hinder the aesthetics of the plots,
and as they are at the limits of our grid we have decided to omit them
from some Figures by trimming the axes. Additionally, stars at the
highest inclinations, i.e edge on disc systems, are often omitted from
the Figures due to their extremely faint magnitudes, meaning they
would not be practically observable.

As discussed in Section \ref{disc_struct} the inner edge location is
correlated with inner edge temperature (albeit differently for the
typical and extreme accretors). As noted by \cite{meyer_1997} this
could lead to a correlation of IR excess with inner edge position. As
expected however, scattered caused by variations in inclination and
disc structure act to remove this correlation for populations.  The
correlation for an individual set of a systems, i.e. all variables
fixed except rotation rate, between IR colour and rotation rate is
preserved. However, as shown in Section \ref{disc_struct} the changes
caused by accretion rate and inclination angle are the most
significant, and act to remove any correlation between rotation rate
and IR colours for populations. In fact, for our grid the simulated
photometry shows no significant correlation between rotation rate and
IR colours. The work of \cite{meyer_1997} studied flat accreting disc,
therefore, as we have increased variation in disc structure by
applying vertical hydrostatic equilibrium it would be interesting to
see if any correlation appears for analytically defined disc
structures. We have run a set of parrallel models using analytical
disc structures and will publish the results in a subsequent
publication. Overall, as found in \cite{walker_2004}, the
dominant scattering effect for BD disc systems in an optical CMD
appears to be caused by accretion rate and inclination. Therefore for
derivation of parameters we explore the scatter in pre-MS BD locii as
a function of accretion rate and inclination.

@@

For the derivation of ages optical CMDs, in particular in \textit{V,
  V-I}, are most often used, and indeed most suitable. Whereas, IR
CMDs, such as a \textit{J, J-K} CMD, are most suitable for mass
derivation \cite[see references and discussions in][ and Section
\ref{derived}]{mayne_2007,mayne_2008}.

The use of pre-MS isochrones for the derivation of single star
parameters is at the present time not proven to be reliable \citep[see
discussion in][]{mayne_2008}. Practically, therefore, median
ages are derived from populations. Subsequently, derived masses are
still unreliable but at least based on a consistent age. This problem
is being addressed by Bell et al (in prep), where $K$ band photometry
and known eclipsing binaries are being used to refine pre-MS
isochrones. In this section we plot the data for our 1 Myr systems
only and explore the resulting scatters caused by the disc presence
and accretion luminosity.

Figures \ref{age_mass_VVI} and \ref{age_mass_JJK} shows CMDs in $M_V$
\& $(V-I)_0$ and $M_J$ \& $(J-K)_0$ resepctively. The left panels of
both figures shows stars classed as typical accretors with accretion
rates of \logmdot=-9, -10 and -11 \& -12, shown as blue, black and red
dots respectively in the top panels. The right panels of Figures
\ref{age_mass_VVI} and \ref{age_mass_JJK} show systems with extreme
accretion rates, with \logmdot=-6, 7 and -8 shown as blue, black and
red dots respectively, in the top right panel. The bottom panels then
show the systems seperated into groups by inclination. These groups
are $\theta \leq 48^{\circ}$ as blue dots (classed as face-on
systems), $\theta> 56$ \& $64^{\circ}$ (classed as the expected
systems, as the expectation value of $cos(\theta)=60^{\circ}$) as
black dots and $\theta \geq 71^{\circ}$ (classed as edge-on systems)
as red dots (for typical, bottom left, and extreme, bottom right
accretion rates). The left panels then show naked stars isochrones
(created from our grid of simulated photometry) for 1 Myrs at
accretion rates of \logmdot=-12 and -9, as solid and dashed lines
respectively (with coverage of 10\% and rotation rates of 5
days). Whereas the left hand panels show naked stars isochrones of
\logmdot=-6 and -9 as solid and dashed black lines respectively (with
coverage of 10\% and rotation rates of 5 days. The solid green line,
in the top panels, shows the 1 Myr isochrone of
\cite{siess_2000} adjusted to a distance of 250pc and an
extinction of $A_V=2$ mag, simulating a background population of CTTS
stars.

\begin{figure*}
  \centering
  \includegraphics[scale=0.6,angle=90]{./Fig/age_mass_VVI.ps}
  \caption{Figure showing CMDs in $M_V$, $(V-I)_0$ for typical
    accretors (\logmdot $\leq$ -9) in the left panels, and extreme
    accretors (\logmdot $>-9$) in the right panels. The black dashed
    and solid lines correspond to naked BD isochrones (from our grid)
    with \logmdot -9 and -12 for the left panels (typical accretors)
    and -9 and -6 in the right panels (extreme accretors). The
    rotation rates and areal coverages are set at 5 days and 10\%
    respectively (changing these has little effect). The top panels
    also include as a solid green line the 1 Myr pre-MS isochrone of
    \citet{siess_2000} adjusted to a distance modulus of 7 and an
    extinction of $A_V=2$, simulating a reddened background
    population. The top panels then seperate the systems by accretion
    rate with \logmdot=-9, -10 and -11 \& -12 in the top left panel,
    and \logmdot=-6, -7 and -8 in the top right panel, shown blue,
    black and red dots respectively for both cases. The lower panels
    then split the systems into groups by inclination, with $\theta
    \leq$ 48$^{\circ}$, $\theta \geq$ 77$^{\circ}$ and $\theta=$ 56,
    64 \& 71$^{\circ}$, plotted as blue, red and black dots
    respectively.\label{age_mass_VVI}}
\end{figure*}

\begin{figure*}
  \centering
  \includegraphics[scale=0.6,angle=90]{./Fig/age_mass_JJK.ps}
  \caption{Figure showing the same data in the same format and with
    the same symbol meanings as Figure \ref{age_mass_VVI}, but for
    $M_J$ and $(J-K)_0$ CMDs. However, for this Figure the \logmdot=-6
    isochrone is not shown (lies significantly blueward of locus of
    data), and the pre-MS isochrone of \citet{siess_2000}, has been
    smoothed.\label{age_mass_JJK}}
\end{figure*}

As can be seen in Figures \ref{age_mass_VVI} and \ref{age_mass_JJK}
current accretion and disc presence in a star and disc systems creates
a scatter in our simulated photometry indicative of a much larger
isochronal age or mass spread. Indeed, for many BDD systems even at
nominal accretion rates of \logmdot= -11 or -12, for our simulations,
the colours of these stars move significantly blueward of the expected
BD locus in a $M_V$, $(V-I)_0$ CMD and redward in a $M_J$, $(J-K)_0$
CMD. As, in the case of typical accretors, the input variables are in
the range expected for a BD population, one could reasonably expect
observed true BD locii to show a similar scatter. It is clear that
even a wide photometric selection would not include all of, even the
negligibly accreting systems at expected inclinations. Whilst the
derivation of masses and ages for these objects will be
difficult. Furthermore, for the higher accretion rates of -9 or -10
\logmdot the movement of the star within the CMD will effectively move
the star into the contamination region expected for background CTTS or
MS stars at a $(V-I)_0$ of $\leq$1.5 and $M_V$ 12-10, and as such the
star would not be included in a photometrically selected BD
sample. The solid green line, in the top panels of Figures
\ref{age_mass_VVI} and \ref{age_mass_JJK}, showing the 1 Myr isochrone
of \cite{siess_2000} at a distance of 250 pc and extinction of $A_V=2$
mags shows that the BDD systems with higher accretion rates could
easily be confused for a background CTTS or MS population. Indeed miss
classification of a BDD system as a CTTS system has already been
revealed in \cite{white_2003}. For the extreme accretors, as shown in
the right panels of Figures \ref{age_mass_VVI} and \ref{age_mass_JJK},
there is little chance of these objects being classed,
photometrically, as BD candidates or being assigned the correct
mass. Therefore, if one attempted to locate a population of BDD
systems with elevated or extreme accretion rates, target selection
would have to be placed at much brighter magnitudes, and bluer for
optical or redder for IR, colours.
